### 2025-03-04 manuscript.Rmd
a whole lot of love for the references

The ´reproducibleRchunks` package addresses the challenge of ensuring computational results in R Markdown documents are automatically testable for successful reproduction. 
This means verifying whether the same script with the same data yields identical results, even on different computers or at later times, with minimal disruption to users' workflows.
This article is intended for readers who use R for statistical analyses and ideally already had some experience with R Markdown. 
We aim to first offer a less technical tutorial outlining a typical use case for the package. 
Following this, we provide more detailed technical background information about the package and various ways to customize its behavior.
We believe that the package can be used effectively without a deep understanding of the technical details. 
However, gaining insight into these details may enhance the user experience and help readers understand the challenges involved in ensuring reproducibility.

@@

```{r rstudioscreen1, fig.cap="Automated testing of variables for reproducibility is initiated by using code chunks of type `reproducibleR` (see yellow highlight) instead of `R` as shown in this example R Markdown snippet. "}
knitr::include_graphics("img/rstudio-screenshot-marker2.png")
```

```{r rstudioscreen2, fig.cap="Snippet from a knitted Markdown file showing the reproducibility report when a document is generated for the first time."}
knitr::include_graphics("img/generation-step1.png")
```

@@

```{r rstudioscreen3, fig.cap="Snippet from a knitted Markdown file showing the reproducibility report when a document is re-generated and a variable is successfully reproduced."}
knitr::include_graphics("img/generation-step2.png")

@@

```{r rstudioscreen4, fig.cap="Snippet from a knitted Markdown file showing the reproducibility report when a document is re-generated and the reproduction of a variable failed.", fig.pos="b"}
knitr::include_graphics("img/generation-step2-failed.png")

@@

- Original computations were executed and stored in the JSON data file. Later, a reproduction attempt is made using the exact same R Markdown file but on a different computer. This is a classic case of non-reproducibility, which often is due to changes in the software packages and the R version the computations in the R Markdown rely upon [@epskamp2019reproducibility; @peikert2021reproducible]. Note that the goal of this package is not to guarantee reproducibility but to allow for automated testing and reporting of reproducibility. To increase the chances of reproducibility in the first place, various solutions exist [@peikert2021reproducible; @renv2022; @chan2023rang; @nagraj2023pracpac] of which a primary aspect is a way to document and recreate the original computational environment.
- Some computations are executed and their original results are stored in a JSON file as planned. Later on, someone modifies the R code in the R Markdown file, such that the results differ and a failure of reproduction is indicated. This change of code between the original computation and the reproduction attempt is caught by the package because fingerprints of the entire code chunk syntax are stored. Users are informed by a warning in the reproducibility report. The package will still try to reproduce each result and give individual reports on successes and failures.
- The metadata files get lost. Without metadata files, no automated reproducibility check can be made. In this case, it may be a good idea to rerun the analysis in a computing environment as close as possible to the original computing environment and store the newly generated metadata files. It is advisable to manually compare the recreated results to some original reference (e.g., a report or published article) for consistency. 
- The `reproducibleRchunks` package is not available anymore. In this case, all R Markdown code chunks of type `reproducibleR` can be renamed to `R` and the manuscript will at least render even though without automated reproducibility checks. If there is a large number of code chunks, the following line of code can be inserted in the Markdown document in the very first code chunk, which will tell knitr to render all `reproducibleR` chunks as regular R chunks without the need to rename the chunks: `R knitr::knit_engines$set(reproducibleR = knitr::knit_engines$get("R")`.
- `R` is not available anymore. Some day `R` may be superseded by a future programming language and/or there will be no port of `R` or the `reproducibleRchunks` package to a future computing platform. Then, the JSON format will likely still allow the possibility to read out the original computational results in that future programming language as the JSON format itself is open and transparent. 
- A future `R` version changes the way objects are serialized (that is, converted from the internal representation to a byte-stream representation, of which the fingerprint is taken). The metadata contains information about the R version used to generate the metadata, so future versions of our package could easily be adapted to this change. As long as the `digest` package is used for generating fingerprints, the serialization version can be fixed to the version that is used at the time of writing (2024) using this command: `options(serializeVersion=2)`.

>>>>

The `reproducibleRchunks` package addresses the challenge of ensuring computational results in R Markdown documents are automatically testable for successful reproduction. 
This means verifying whether the same script with the same data yields identical results, even on different computers or at later times, with minimal disruption to users' workflows.
This article is intended for readers who use R for statistical analyses and ideally already had some experience with R Markdown. 
We aim to first offer a less technical tutorial outlining a typical use case for the package. 
Following this, we provide more detailed technical background information about the package and various ways to customize its behavior.
We believe that the package can be used effectively without a deep understanding of the technical details. 
However, gaining insight into these details may enhance the user experience and help readers understand the challenges involved in ensuring reproducibility.

@@

```{r rstudioscreen1, fig.cap="Automated testing of variables for reproducibility is initiated by using code chunks of type `reproducibleR` (see yellow highlight)"}
knitr::include_graphics("img/rstudio-screenshot-marker2.png")
```

```{r rstudioscreen2, fig.cap="Snippet from a knitted Markdown file showing the reproducibility report of a code chunk when a document is generated for the first time."}
knitr::include_graphics("img/generation-step1.png")
```

@@

```{r rstudioscreen3, fig.cap="Snippet from a knitted Markdown file showing the reproducibility report of a code chunk when a document is re-generated and a result is successfully reproduced."}
knitr::include_graphics("img/generation-step2.png")

@@

```{r rstudioscreen4, fig.cap="Snippet from a knitted Markdown file showing the reproducibility report when a document is re-generated and the reproduction of a result failed.", fig.pos="b"}
knitr::include_graphics("img/generation-step2-failed.png")

@@

- Original computations were executed and stored in the JSON data file. Later, a reproduction attempt is made using the exact same R Markdown file but on a different computer. This is a classic case of non-reproducibility, which often is due to changes in the software packages and the R version the computations in the R Markdown rely upon [@epskamp2019reproducibility; @peikert2021reproducible]. Note that the goal of this package is not to guarantee reproducibility but to allow for automated testing and reporting of reproducibility. To increase the chances of reproducibility in the first place, various solutions exist [@peikert2021reproducible; @renv2022; @chan2023rang; @nagraj2023pracpac] of which a primary aspect is a way to document and recreate the original computational environment.
- Some computations are executed and their original results are stored in a JSON file as planned. Later on, someone modifies the R code in the R Markdown file, such that the results differ and a failure of reproduction is indicated. This change of code between the original computation and the reproduction attempt is caught by the package because fingerprints of the entire code chunk syntax are stored. Users are informed by a warning in the reproducibility report. The package will still try to reproduce each result and give individual reports on successes and failures.
- The metadata files get lost. Without metadata files, no automated reproducibility check can be made. In this case, it may be a good idea to rerun the analysis in a computing environment as close as possible to the original computing environment and store the newly generated metadata files. It is advisable to manually compare the recreated results to some original reference (e.g., a report or published article) for consistency. 
- The `reproducibleRchunks` package is not available anymore. In this case, all R Markdown code chunks of type `reproducibleR` can be renamed to `R` and the manuscript will at least render even though without automated reproducibility checks. If there is a large number of code chunks, the following line of code can be inserted in the Markdown document in the very first code chunk, which will tell knitr to render all `reproducibleR` chunks as regular R chunks without the need to rename the chunks: `R knitr::knit_engines$set(reproducibleR = knitr::knit_engines$get("R")`.
- `R` is not available anymore. Some day `R` may be superseded by a future programming language and/or there will be no port of `R` or the `reproducibleRchunks` package to a future computing platform. Then, the JSON format will likely still allow the possibility to read out the original computational results in that future programming language as the JSON format itself is open and transparent. 
- A future `R` version changes the way objects are serialized (that is, converted from the internal representation to a byte-stream representation, of which the fingerprint is taken). The metadata contains information about the R version used to generate the metadata, so future versions of our package could easily be adapted to this change. As long as the `digest` package is used for generating fingerprints, the serialization version can be fixed to the version that is used at the time of writing (2025) using this command: `options(serializeVersion=2)`.
